Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-02-28 paper1_jiturra.Rmd
up

As social relations within the occupational structure serve as the foundation for class analysis, widening our examination to other dimensions of social life, such as family dynamics, the development of friendships, and diverse social activities, can significantly enrich the scope of class analysis, particularly concerning its relevance in political attitudes formation. In principle, social stratification scholars have traditionally considered _differentiation_ as the observable patterns of association given by the intersection of status and group membership, which can be understood as relational networks between positions [@blau_inequality_1977]. Empirically, homophily is one of the most significant findings in network research, understood as the higher likelihood of tie formation between people of similar characteristics [@mcpherson_birds_2001; p.416] around demographic characteristics that trace friendship and family ties such as age and gender, as well as the segregated social environments alongside ethnicity and socioeconomic status [@bargsted_pautas_2020; @plaza_juntos_2022,@lazarsfeld_friendship_1954]. In addition, psychosocial factors such as socialization preferences and group dynamics also play a role in segregation processes, considering that isolated social environments decrease the interchange of world views, reinforcing in-group selectivity and value similarity [@visser_attitudes_2004;@homans_human_1951]. Nonetheless, without denying the relevance of socialization preferences, segregation has been predominantly explained through social differentiation in social activities as foci of contact opportunities that ultimately consolidate social networks [@feld_focused_1981; @mcpherson_network_2019]. From this scope, we argue that class relations represent not only resource-based distinctions but also patterns of sociability and exchange observed in the differentiation of social ties.

>>>>

As social relations within the occupational structure serve as the foundation for class analysis, widening our examination to other dimensions of social life, such as family dynamics, the development of friendships, and diverse social activities, can significantly enrich the scope of class analysis, particularly concerning its relevance in political attitudes formation. In principle, social stratification scholars have traditionally considered _differentiation_ as the observable patterns of association given by the intersection of status and group membership, which can be understood as relational networks between positions [@blau_inequality_1977]. Empirically, homophily is one of the most significant findings in network research, understood as the higher likelihood of tie formation between people of similar characteristics [@mcpherson_birds_2001, p.416] around demographic characteristics that trace friendship and family ties such as age and gender, as well as the segregated social environments alongside ethnicity and socioeconomic status [@bargsted_pautas_2020;@plaza_juntos_2022;@lazarsfeld_friendship_1954]. In addition, psychosocial factors such as socialization preferences and group dynamics also play a role in segregation processes, considering that isolated social environments decrease the interchange of world views, reinforcing in-group selectivity and value similarity [@visser_attitudes_2004;@homans_human_1951]. Nonetheless, without denying the relevance of socialization preferences, segregation has been predominantly explained through social differentiation in social activities as foci of contact opportunities that ultimately consolidate social networks [@feld_focused_1981; @mcpherson_network_2019]. From this scope, we argue that class relations represent not only resource-based distinctions but also patterns of sociability and exchange observed in the differentiation of social ties.
